---
title: "McCammon Exercise 8"
output:
  html_document:
    df_print: paged
---

### 1. 
#### Create a function that will compute the total score for respondents on a Likert scale. The function accepts a data frame, a numeric with the scale used, and a numeric vector indicating the item number that should be reversed. 

```{r echo=TRUE}
likert.score <- function(df, scale, item) {
  reverse.score <- df[,item] <- (scale+1) - df[,item]
  total.score <- apply(df,1, sum)
  personality.total <- cbind(df,total.score)
  return(personality.total)
}
```

### 2. 
#### Create a CSV file that contains data from a hypothetical survey with eight Likert items using a six-point scale with three of these being revese scored. Write a script that reads in your data, sends it to your function for scoring then writes the results to a csv file that now contains the original data with the score added 

```{r include=FALSE}
library(here)
```
```{r include=FALSE}
personality <- read.csv(here("Data","Personality.csv"))
```
```{r echo=FALSE}
personality.total <-likert.score(personality, 6, c(3,4,8))
personality.total
```
```{r include=FALSE}
write.csv(personality.total,file = "Personality (with total score).csv",
          row.names = FALSE) 
```

### 3. 
#### Use the hsb2 data and ggplot2 to construct a single scatterplot showing the relationship of science scores (response) to math scores (explanatory) by gender

```{r include=FALSE}
library(tidyverse)
```
```{r include=FALSE}
library(here)
```
```{r include=FALSE}
hsb2 <- read_csv(here("Data","hsb2.csv"))
```
```{r include=FALSE}
hsb2$female <- factor(hsb2$female, labels = c("female","male"), ordered = FALSE)
```
```{r echo=FALSE}
ggplot(data = hsb2) +
  geom_point(mapping = aes(x=math, y = science, color = female))
```

### 4. 
#### Show the relationship of science scores to math scores by gender, but this time facet on gender to show separate scatterplots

```{r echo=FALSE}
ggplot(data = hsb2) +
  geom_point(mapping = aes(x=math, y = science)) +
  facet_wrap(~female)
```

### 5. 
#### Using ggplot2 and hsb2 data, create a graph using a geom that was not mentioned in the lecture. Describe what the graph illustrates
*This graph displays the scatterplot with the regression line for the relationship between math and science scores by gender, we see that female scores vary more than males*

```{r echo=FALSE}
ggplot(data = hsb2, mapping = aes(x = math, y = science, col = female))+
    geom_point() +
    geom_abline()+
    facet_wrap(~female)
```
